---
title: 'Text mining'
date: 'created on 22 November 2020 and updated `r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(here)


# For text mining:
library(pdftools)
library(tidytext)
library(textdata) 
library(ggwordcloud)
library(lubridate)
library(tibble)



# Note - Before lab:
# Attach tidytext and textdata packages
# Run: get_sentiments(lexicon = "nrc")
# Should be prompted to install lexicon - choose yes!
# Run: get_sentiments(lexicon = "afinn")
# Should be prompted to install lexicon - choose yes!

```


### Get the lettet:
```{r get-document}
letter1_path <- here("data","brev_Rosenkrantz-02.pdf")
library(tidytext)
letter1_text <- pdf_text(letter1_path)
letter1_text
```

The first page of the pdf-file is only regarding administrative information from the distributor regarding the document. Therefore we only need page 2 for the analysis, which contains the letter. 

```{r single-page}
letter1_p2 <- letter1_text[2]
letter1_p2
```

### Some wrangling:

- Split up pages into separate lines (separated by `\n`) using `stringr::str_split()`
- Unnest into regular columns using `tidyr::unnest()`
- Remove leading/trailing white space with `stringr::str_trim()`

```{r split-lines}
letter1_df <- data.frame(letter1_p2) %>% 
  mutate(text_full = str_split(letter1_p2, pattern = '\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 
letter1_df
# Why '\\n' instead of '\n'? Because some symbols (e.g. \, *) need to be called literally with a starting \ to escape the regular expression. For example, \\a for a string actually contains \a. So the string that represents the regular expression '\n' is actually '\\n'.
# Although, this time round, it is working for me with \n alone. Wonders never cease.

# More information: https://cran.r-project.org/web/packages/stringr/vignettes/regular-expressions.html

```

Now each line, on each page, is its own row, with extra starting & trailing spaces removed. 

### Get the tokens (individual words) in tidy format

Use `tidytext::unnest_tokens()` (which pulls from the `tokenizer`) package, to split columns into tokens. We are interested in *words*, so that's the token we'll use:

```{r tokenize}
letter1_tokens <- letter1_df %>% 
  unnest_tokens(word, text_full)
letter1_tokens


# Each word has its own row!
```

Let's count the words!
```{r count-words}
letter1_wc <- letter1_tokens %>% 
  count(word) %>% 
  arrange(-n)
letter1_wc
```

OK...so we notice that a whole bunch of things show up frequently that we might not be interested in ("a", "the", "and", etc.). These are called *stop words*. Let's remove them. 

### Remove stop words:
```{r}
stoplist <- readLines("stoplist.txt")
stoplist
stoplist <- trimws(stoplist)
```

With help from Chatgpt we formed a script for inserting our own stoplist customized for the gothic language from 1500
```{r}
# Install and load necessary libraries if you haven't already
install.packages("dplyr")
library(dplyr)

# Convert stoplist into a data frame
stoplist_df <- data.frame(word = stoplist)

# Use anti_join to remove stopwords from the text data
letter1_stop <- anti_join(letter1_wc, stoplist_df, by = "word")

# View the cleaned text data (with stopwords removed)
print(letter1_stop)
```


We will *remove* stop words using `tidyr::anti_join()`:



Now check the counts again: 
```{r count-words2}
letter1_swc <- letter1_stop %>% 
  count(word) %>% 
  arrange(-n)
letter1_swc
```

What if we want to get rid of all the numbers (non-text) in `ipcc_stop`?
```{r skip-numbers}
# This code will filter out numbers by asking:
# If you convert to as.numeric, is it NA (meaning those words)?
# If it IS NA (is.na), then keep it (so all words are kept)
# Anything that is converted to a number is removed

letter1_no_numeric <- letter1_stop %>% 
  filter(is.na(as.numeric(word)))

letter1_no_numeric
```

### A word cloud of the letter words (non-numeric)

See more: https://cran.r-project.org/web/packages/ggwordcloud/vignettes/ggwordcloud.html

```{r wordcloud-prep}
# There are almost 2000 unique words 
length(unique(letter1_no_numeric$word)) # shows the unique words and count them

# We probably don't want to include them all in a word cloud. Let's filter to only include the top 100 most frequent?
letter1_top150 <- letter1_no_numeric %>% 
  count(word) %>% 
  arrange(-n) %>% 
  slice(1:150) # or head

letter1_top150
```

```{r wordcloud}
letter1_cloud <- ggplot(data = letter1_top150, aes(label = word)) +
  geom_text_wordcloud() +
  theme_minimal()

letter1_cloud
```

That's underwhelming. Let's customize it a bit:
```{r wordcloud-pro}
ggplot(data = letter1_top150, aes(label = word, size = n)) +
  geom_text_wordcloud_area(aes(color = n), shape = "diamond") +
  scale_size_area(max_size = 8) +
  scale_color_gradientn(colors = c("darkgreen","blue","red")) +
  theme_minimal()
```
```{r}
#Indsat ved vejledning fra chatgpt til at kunne forbinde de indsatte filer til at kunne anvende funktionen "analyze()"
library(pdftools)

analyze <- function(file_path) {
  text <- pdf_text(file_path)
  if (length(text) >= 2) {
    cat(text[2])
  } else {
    cat("PDF has fewer than 2 pages.\n") #ønsker kun 2. side i hver dokument
  }
}
```



```{r}
#Følger guiden fra software carpentry 
list.files(path = "data", pattern = "pdf")
```
```{r}
#Følger guiden fra software carpentry 
list.files(path = "data", pattern = "brev_Rosenkrantz", full.names = TRUE)
```
```{r}
#Følger guiden fra software carpentry 
filenames <- list.files(path = "data",  
                        pattern = "brev_Rosenkrantz-[0-9]{2}.pdf",
                        full.names = TRUE)
for (f in filenames) {
  print(f)
  analyze(f)
}
```


```{r}
#Kombinere det med tidytext
analyze <- function(file_path) {
  text <- pdf_text(file_path)
  
  if (length(text) < 2) {
    return(tibble(file = basename(file_path), word = NA))
  }

  second_page <- text[2]
  #ønsker kun 2. side på hver dokument
  
  tibble(file = basename(file_path), text = second_page) %>%
    unnest_tokens(word, text)
}
```


```{r}
filenames <- list.files(path = "data",  
                        pattern = "brev_Rosenkrantz-[0-9]{2}.pdf",
                        full.names = TRUE)

# Use lapply to process all files and bind results
all_words <- lapply(filenames, analyze) %>%
  bind_rows()
```

```{r}
all_words
#indeholder kun for de tre første pdf - måske får mange ord? brug for troubleshooting
```
```{r}
#Tjek op på at alle 9 filer er tilstede 
print(length(filenames)) 
print(basename(filenames))
```

```{r}
#Fjerner stopord 
my_stops <- readLines("stoplist.txt")
my_stops_df <- tibble(word = my_stops)

all_words_clean <- all_words %>%
  filter(!is.na(word)) %>%
  anti_join(my_stops_df, by = "word")

all_words_clean_no_numeric <- all_words_clean %>% 
  filter(is.na(as.numeric(word))) %>% 
  count(word, sort=TRUE) 

all_words_clean_no_numeric %>% 
  arrange(word)
  
all_words_clean_no_numeric


```

```{r}
view(all_words_clean_no_numeric)
```

```{r}

unique_words <- all_words_clean_no_numeric %>%
  count(word, sort = TRUE) %>%
  filter(n == 1)

unique_words

```

```{r}
all_words <- all_words %>%
  mutate(year = as.numeric(year),
         year = 1500 + year)

```


```{r}
all_words %>% 
  count(word, file, sort = TRUE)

```



